High-Performance Primitives for Distributed Memory Architectures using Asynchronous Events

### 2012-11-14 doc/llr-pldi2013/impl.tex
last-minute edits, spell check

The heterogeneous implementation
models a cluster with CPUs and GPUs as having two kinds of processors 
and four kinds of memory.  Every CPU core is presented as a different {\tt CPU} processor
and every GPU is a {\tt GPU} processor.  This matches the scheduling granularity available
in the Pthreads and CUDA APIs respectively.  The first kind of memory is the 
system memory that is accessible by every CPU core on a given node.  The second is the 
framebuffer memory on each GPU (and accessible only by that GPU).  The third kind
of memory is system memory on a node that has been made accessible to the GPU(s) on
that node as well as the CPU cores, commonly referred to as {\em zero-copy memory}.
The final kind of memory is the portion of system memory on each node that has been registered with
the GASNet runtime to allow {\em remote memory access} (RMA) by other nodes in the cluster.
We refer to this memory as {\em GASNet memory}.
%% For every node in the system
%% there is a memory corresponding to the DRAM associated with that node.  For
%% every GPU in the system there are two memories corresponding to the framebuffer
%% memory and the zero-copy memory associated with that GPU \footnote{Zero-copy 
%% memory contains pages that are mapped in both framebuffer and node memory with
%% coherence maintained by the PCI-E protocol.}. The last memory is 
%% a global GASNet memory that represents pages that have been registered 
%% with all nodes in the system for supporting remote memory accesses (RMAs).
%% The GASNet memory provides the illusion of global memory.  CPU processors
%% can directly access their node memory, all zero-copy memories for GPUs on
%% their node, and the GASNet memory.  GPU processors can only access their
%% framebuffer memory and their zero-copy memory.  Copies are permitted
%% between all pairs of memory except between any framebuffer memory and
%% the GASNet memory.

In addition to the RMA capabilities, GASNet provides {\em active messages} for 
inter-node communication.  Implementation of all runtime features requiring communication
rely heavily on active messages.  Active messages consist of a command and payload that
are sent by one node to another without any previous coordination.  Upon arrival
at a destination node, a handler routine is invoked to process the message.
%The handler may issue a response to the sender, but this is optional, and avoided
%as much as possible in our implementation due to the latencies involved in waiting
%for a response.  
%One example of a case where we use active messages involves the {\tt spawn} call on a {\tt Processor}
%object.  If the processor is not on the same node as the requestor, the information
%about the task (e.g. the task ID and arguments, the actual target {\tt Processor},
%the prerequisite event (if any), and the completion event ID) 
%for the task is sent in an
%active message to the node containing the target processor.  The recipient node
%unpacks the message, and either places the task on the ready queue for the target
%processor or marks it as being deferred until the prerequisite event has triggered.
%No response is sent to the original requestor.  The requestor already knows the event
%that corresponds to the completion of this task (since it was provided in the active
%message) - there is no utility in knowing exactly when the active message itself has
%been processed.

@@

%At event creation time, the owning node allocates a data structure to track the state of the event
%(i.e. triggered or not) as well as record a list of the operations that are known to be dependent
%on the event (e.g. initiation of a copy operation, placing a task on a processor's ready queue, waking
%up a waiting task).  However, the owning node only keeps a list of the dependent operations from the same
%node.  Every other node also allocates a corresponding data structure (the first time the event
%is referenced) to remember if the event has already triggered and to record its local dependent operations.
%Arbitrarily many dependent operations from a
%single node are aggregated into a single {\em event subscription} active message that is sent to the
%owning node.  The owning node keeps a bitmask of which other nodes have subscribed to the event, and when the
%event finally does trigger, a single {\em event trigger} active message is sent to each subscribing node,
%which notes that the event has triggered (in case queries come after the trigger) and executes the list
%of local operations that were dependent on that event.  (In the case that the triggering of the event has
%occurred before the reception of an event subscription, a trigger message is sent immediatel to the 
%new subscriber.)
When a new event is created, the owning node allocates a data structure to track the state of
the event ({\em triggered} or {\em untriggered}) as well as to record a list of dependent operations
(e.g. copy operation or task launch) from the same node called {\em waiters}.
The first operation dependent on an event
from a remote node allocates the same data structure on the remote node.  A single
{\em event subscription} active message is then sent from the remote node to the owning node indicating
that the remote node should be informed when the event triggers.  Arbitrarily many dependent operations on
the remote node can then be added to the list of waiters without any additional communication.
When the event does trigger, the owner node notifies all local waiters and
sends a single {\em event trigger} active message to each node from which it has seen an event subscription
to notify remote waiters.
In the case where an event triggers while a subscription message is in flight from a remote node, the owner node will
immediately respond with a trigger message.

@@

Like events, deferred locks (henceforth called simply locks) are
created on demand by the node of the creation request. 
Lock handles are also statically divided across the nodes,
allowing creation to be done without inter-node communication.
Whereas event ownership is static, locks may migrate; the {\em
  creating} node is the initial owner, but ownership can be
transferred to other nodes.

@@

When a lock request is made, an event is created to track when the grant occurs.  The current node then
examines its copy of the lock data structure to determine if it is the owner.
%(creating it if this the first operation for that lock on that node).  
If the current node is the owner and the lock isn't
held, the lock is granted immediately and the event is triggered.  If the lock is held, the event is added
to the list of local waiters.  Note that the event associated with
the lock request is the only data that must be stored.
%(This is the only thing that is stored.  The actual ``requestor'' of the lock
%is implicitly captured in which operations are dependent on the lock grant event.)  
If the current node isn't the owner, a {\em lock request} active message is
sent to the most recently known owner.  If the receiver of a lock request message is no
longer the owner, it forwards the message on to the node it has recorded as the owner.
%(If that information proves to be stale, the request is forwarded
%by the receiving node, eventually catching up to the current owner.)  
If the current owner's status shows
the lock is currently held, the bit for the requesting node is set in the remote waiters bitmask.  If the lock is
not held, the ownership of the lock is given to the requesting node via a {\em lock transfer} active
message, which includes the bitmask of remote waiters and an up-to-date copy 
of the lock's payload.  The inclusion of the payload in the active message is the reason for the 
4KB size limit specified on payloads in Section~\ref{subsec:locks}.

>>>>

The heterogeneous implementation
models a cluster with CPUs and GPUs as having two kinds of processors 
and four kinds of memory.  Every CPU core is presented as a different {\tt CPU} processor
and every GPU is a {\tt GPU} processor.  This matches the scheduling granularity available
in the Pthreads and CUDA APIs respectively.  The first kind of memory is the 
system memory that is accessible by every CPU core on a given node.  The second is the 
framebuffer memory on each GPU (and accessible only by that GPU).  The third kind
of memory is system memory on a node that has been made accessible to the GPU(s) on
that node as well as the CPU cores, commonly referred to as {\em zero-copy memory}.
The final kind of memory is the portion of system memory on each node that has been registered with
the GASNet runtime to allow {\em remote direct memory access} (RDMA) by other nodes in the cluster.
We refer to this memory as {\em GASNet memory}.
%% For every node in the system
%% there is a memory corresponding to the DRAM associated with that node.  For
%% every GPU in the system there are two memories corresponding to the framebuffer
%% memory and the zero-copy memory associated with that GPU \footnote{Zero-copy 
%% memory contains pages that are mapped in both framebuffer and node memory with
%% coherence maintained by the PCI-E protocol.}. The last memory is 
%% a global GASNet memory that represents pages that have been registered 
%% with all nodes in the system for supporting remote memory accesses (RDMAs).
%% The GASNet memory provides the illusion of global memory.  CPU processors
%% can directly access their node memory, all zero-copy memories for GPUs on
%% their node, and the GASNet memory.  GPU processors can only access their
%% framebuffer memory and their zero-copy memory.  Copies are permitted
%% between all pairs of memory except between any framebuffer memory and
%% the GASNet memory.

In addition to the RDMA capabilities, GASNet provides {\em active messages} for 
inter-node communication.  Implementation of all runtime features requiring communication
rely heavily on active messages.  Active messages consist of a command and payload that
are sent by one node to another without any previous coordination.  Upon arrival
at a destination node, a handler routine is invoked to process the message.
%The handler may issue a response to the sender, but this is optional, and avoided
%as much as possible in our implementation due to the latencies involved in waiting
%for a response.  
%One example of a case where we use active messages involves the {\tt spawn} call on a {\tt Processor}
%object.  If the processor is not on the same node as the requestor, the information
%about the task (e.g. the task ID and arguments, the actual target {\tt Processor},
%the prerequisite event (if any), and the completion event ID) 
%for the task is sent in an
%active message to the node containing the target processor.  The recipient node
%unpacks the message, and either places the task on the ready queue for the target
%processor or marks it as being deferred until the prerequisite event has triggered.
%No response is sent to the original requestor.  The requestor already knows the event
%that corresponds to the completion of this task (since it was provided in the active
%message) - there is no utility in knowing exactly when the active message itself has
%been processed.

@@

%At event creation time, the owning node allocates a data structure to track the state of the event
%(i.e. triggered or not) as well as record a list of the operations that are known to be dependent
%on the event (e.g. initiation of a copy operation, placing a task on a processor's ready queue, waking
%up a waiting task).  However, the owning node only keeps a list of the dependent operations from the same
%node.  Every other node also allocates a corresponding data structure (the first time the event
%is referenced) to remember if the event has already triggered and to record its local dependent operations.
%Arbitrarily many dependent operations from a
%single node are aggregated into a single {\em event subscription} active message that is sent to the
%owning node.  The owning node keeps a bitmask of which other nodes have subscribed to the event, and when the
%event finally does trigger, a single {\em event trigger} active message is sent to each subscribing node,
%which notes that the event has triggered (in case queries come after the trigger) and executes the list
%of local operations that were dependent on that event.  (In the case that the triggering of the event has
%occurred before the reception of an event subscription, a trigger message is sent immediatel to the 
%new subscriber.)
When a new event is created, the owning node allocates a data structure to track the state of
the event ({\em triggered} or {\em untriggered}) as well as to record a list of dependent operations
(e.g. copy operation or task launch) from the same node called {\em waiters}.
The first operation dependent on an event
from a remote node allocates the same data structure on the remote node.  An
{\em event subscription} active message is then sent to the owning node indicating
that the remote node should be informed when the event triggers.  Arbitrarily many dependent operations on
the subscribing node can be added to the list of waiters without any additional communication.
When the event triggers, the owner node notifies all local waiters and
sends an {\em event trigger} active message to each node from which it has seen an event subscription
to notify remote waiters.
In the case where an event triggers while a subscription message is in flight from a remote node, the owner
node will immediately respond with a trigger message upon receipt of the subscription message.

@@

Like events, deferred locks (henceforth called simply locks) are
created on demand, using a space of handles that has been divided
across the nodes.  As with events, lock creation requires no communication
and the handle alone is sufficient for any node to determine the creator
of a lock.  However, whereas event ownership is static, locks may migrate; the
creating node is the initial owner, but ownership can be
transferred to other nodes.

@@

Each time a lock request is made, a new event is created to track when the grant occurs.  The current node then
examines its copy of the lock data structure to determine if it is the owner.
%(creating it if this the first operation for that lock on that node).  
If the current node is the owner and the lock isn't
held, the lock is granted immediately and the event is triggered.  If the lock is held, the event is added
to the list of local waiters.  Note that the event associated with
the lock request is the only data that must be stored.
%(This is the only thing that is stored.  The actual ``requestor'' of the lock
%is implicitly captured in which operations are dependent on the lock grant event.)  
If the current node isn't the owner, a {\em lock request} active message is
sent to the most recently known owner.  If the receiver of a lock request message is no
longer the owner, it forwards the message on to the node it has recorded as the owner.
%(If that information proves to be stale, the request is forwarded
%by the receiving node, eventually catching up to the current owner.)  
If the current owner's status shows
the lock is currently held, the bit for the requesting node is set in the remote waiters bitmask.  If the lock is
not held, the ownership of the lock is given to the requesting node via a {\em lock transfer} active
message, which includes the bitmask of remote waiters and an up-to-date copy 
of the lock's payload.  The inclusion of the payload in the active message is the reason for the 
4KB size limit specified on payloads in Section~\ref{subsec:locks}.

### 2012-11-14 doc/llr-pldi2013/interface.tex
last-minute edits, spell check

\subsection{Events}
\label{subsec:events}
Events are the primary mechanism for describing dependences between operations in our system.
Lines 1-13 of Figure~\ref{fig:runtimeapi} show the interface for events.  An instance of the {\tt Event} type 
names a unique event in the system.  {\tt NO\_EVENT} (line 3)
is a special instance
of an event that by definition has always triggered.  The event interface
supports testing whether an event has triggered (line 5) and waiting on
an event to trigger (line 6), but the preferred
use of events is passing them as preconditions for other operations.  The client can use the
{\tt merge\_events} call (line 7) to create an event that triggers only when every
event in a given set of events has triggered.  A code example using events follows in
Section~\ref{subsec:procs}.
%corresponding to the conjunction of a set of events.

@@

\begin{figure}
\begin{lrbox}{\mylistingbox}
\begin{lstlisting}
void lock_example(const void *args, size_t arglen, Processor p) {
  // Unpack lock, input events, target processors from arguments
  Lock needed = ...
  // Acquire lock, launch first subtask, release lock
  Event lock_event_1 = needed.lock(prev_event_1);
  Event task_event_1 = proc1.spawn(SUB, args_1, sizeof(args_1), lock_event_1);
  needed.unlock(task_event_1);
  // same for second subtask
  Event lock_event_2 = needed.lock(prev_event_2);
  Event task_event_2 = proc2.spawn(SUB, args_2, sizeof(args_2), lock_event_2);
  needed.unlock(task_event_2);
}
\end{lstlisting}
\end{lrbox}
\subfigure{\usebox{\mylistingbox}}\\

\centering
\scalebox{0.8}{
\begin{tikzpicture}
  \node (l1) at (2,1.95) [shape=rectangle,draw] {\begin{tabular}{c}{\bf lock}\\needed\end{tabular}};
  \node (t1) at (4.5,1.95) [shape=rectangle,draw] {\begin{tabular}{c}{\bf run}\\SUB(args\_1)\end{tabular}};
  \node (u1) at (7,1.95) [shape=rectangle,draw] {\begin{tabular}{c}{\bf unlock}\\needed\end{tabular}};

  \draw [->] (0,1.95) to (l1.west);
  \draw [->] (l1.east) to (t1.west);
  \draw [->] (t1.east) to (u1.west);

  \node (l2) at (2,0.25) [shape=rectangle,draw] {\begin{tabular}{c}{\bf lock}\\needed\end{tabular}};
  \node (t2) at (4.5,0.25) [shape=rectangle,draw] {\begin{tabular}{c}{\bf run}\\SUB(args\_2)\end{tabular}};
  \node (u2) at (7,0.25) [shape=rectangle,draw] {\begin{tabular}{c}{\bf unlock}\\needed\end{tabular}};

  \draw [->] (0,0.25) to (l2.west);
  \draw [->] (l2.east) to (t2.west);
  \draw [->] (t2.east) to (u2.west);

@@

The interface for deferred locks is shown in lines 30-38 of Figure~\ref{fig:runtimeapi}.
Unlike traditional blocking locks, the {\tt lock} method (line 32) doesn't
wait if the lock is in use but instead always returns immediately with an event that will be triggered
when the lock has been acquired.  Like a processor's spawn method, both the {\tt lock} 
and {\tt unlock} methods take an optional event paramenter as a precondition (lines 32-33).

An important difference between deferred locks and blocking locks is that the processor
that requests the lock doesn't have to be the one that uses it.  A common
convention in writing code with deferred locks is to acquire locks on behalf of a task being
launched.  Figure~\ref{fig:lockevents} illustrates this with a simple example.  The {\tt lock\_example}
function needs to launch two subtasks on two other processors, each with exclusive access to some resource protected by a
lock named {\tt mutex}.  For each subtask, it first requests the lock on the subtask's behalf, making the
subtask wait until the lock has been acquired.  The parent task also requests that the runtime automatically
release the lock after each subtask has completed.  In the event dependency graph, solid arrows represent
the explicit events used in the code, while the dotted line represents the dynamic ordering that results
from the mutual exclusion guarantee of the lock.  (In this example, the second subtask is granted the lock first.)

@@

\subsection{Physical Regions}
\label{subsec:phyreg}
%In distributed machines with discrete memories, data movement often consists of more
%than a simple copy operation.  Many applications perform operations and transformations on
%their data in conjunction with data movement for higher performance.  One common example is 
%that applications accumulate reduction operations in a buffer and then, as part of a copy operation,
%the reduction operations are applied to a destination buffer on the receiving side of the copy.  
%To support these kinds of conjoined operations in an asynchronous environment, our interface 
%must be aware of the structure of data so it can perform accompanying operations with 
%data movement operations.  To describe the structure of data our interface uses {\em physical regions.}
Current APIs for moving data in machines with discrete memories only support untyped
copies of a buffer of data and force the programmer to work with pointers in a variety of address
spaces.  This restriction requires programmers to manage both the layout of
data and the marshalling of data for communication.  {\em Physical regions} are a
mechanism for decoupling the placement of data from its layout.  Physical regions allow portability
of ``pointers'' as data is moved around the memory hierarchy, and enable multiple data
layouts, such as specialized layouts for buffering individual reduction operations.

@@

%The data held inside a physical region is accessed by another kind of object called an
%{\em accessor} (not shown).  Accessors support read, write, and reduction operations
%to elements inside the physical region.  Accessors can be specialized for the case where
%all memory is known to be directly accessible to the executing processor, which allows 
%for direct references to elements.  In cases where not all memory is directly accessable, 
%accessors supply the level of indirection necessary to convert operations into remote 
%memory operations (RMAs) if supported by the underlying hardware.  
%An example of this is described in more detail in Section~\ref{sec:impl}.

>>>>

\subsection{Events}
\label{subsec:events}
Events are the primary mechanism for describing dependences between operations in our system.
Lines 1-13 of Figure~\ref{fig:runtimeapi} show the interface for events.  An instance of the {\tt Event} type 
names a unique event in the system.  {\tt NO\_EVENT} (line 3)
is a special instance
of an event that by definition has always triggered.  The event interface
supports testing whether an event has triggered (line 5) and waiting on
an event to trigger (line 6), but the preferred
use of events is passing them as preconditions for other operations.  The client can use the
{\tt merge\_events} call (line 7) to create an aggregate event that triggers once every
event in a given set of events has triggered.  A code example using events follows in
Section~\ref{subsec:procs}.
%corresponding to the conjunction of a set of events.

@@

\begin{figure}
\begin{lrbox}{\mylistingbox}
\begin{lstlisting}
void lock_example(const void *args, size_t arglen, Processor p) {
  // Unpack lock, input events, target processors from arguments
  Lock mutex = ...
  // Acquire lock, launch first subtask, release lock
  Event lock_event_1 = mutex.lock(prev_event_1);
  Event task_event_1 = proc1.spawn(SUB, args_1, sizeof(args_1), lock_event_1);
  mutex.unlock(task_event_1);
  // same for second subtask
  Event lock_event_2 = mutex.lock(prev_event_2);
  Event task_event_2 = proc2.spawn(SUB, args_2, sizeof(args_2), lock_event_2);
  mutex.unlock(task_event_2);
}
\end{lstlisting}
\end{lrbox}
\subfigure{\usebox{\mylistingbox}}\\

\centering
\scalebox{0.8}{
\begin{tikzpicture}
  \node (l1) at (2,1.95) [shape=rectangle,draw] {\begin{tabular}{c}{\bf lock}\\mutex\end{tabular}};
  \node (t1) at (4.5,1.95) [shape=rectangle,draw] {\begin{tabular}{c}{\bf run}\\SUB(args\_1)\end{tabular}};
  \node (u1) at (7,1.95) [shape=rectangle,draw] {\begin{tabular}{c}{\bf unlock}\\mutex\end{tabular}};

  \node[auto] (x1) at (-0.3,2.15) {prev\_event\_1};
  %\draw [->] (x1) to (l1.west);
  \draw [->] (0,1.95) to (l1.west);
  \draw [->] (l1.east) to (t1.west);
  \draw [->] (t1.east) to (u1.west);

  \node (l2) at (2,0.25) [shape=rectangle,draw] {\begin{tabular}{c}{\bf lock}\\mutex\end{tabular}};
  \node (t2) at (4.5,0.25) [shape=rectangle,draw] {\begin{tabular}{c}{\bf run}\\SUB(args\_2)\end{tabular}};
  \node (u2) at (7,0.25) [shape=rectangle,draw] {\begin{tabular}{c}{\bf unlock}\\mutex\end{tabular}};

  \node[auto] (x2) at (-0.3,0.45) {prev\_event\_2};
  \draw [->] (0,0.25) to (l2.west);
  \draw [->] (l2.east) to (t2.west);
  \draw [->] (t2.east) to (u2.west);

@@

The interface for deferred locks is shown in lines 30-38 of Figure~\ref{fig:runtimeapi}.
Unlike traditional blocking locks, the {\tt lock} method (line 32) doesn't
wait if the lock is in use but instead always returns immediately with an event that will be triggered
when the lock has been acquired.  Like a processor's spawn method, both the {\tt lock} 
and {\tt unlock} methods take an optional event parameter as a precondition (lines 32-33).

An important difference between deferred locks and blocking locks is that the processor
that requests the lock doesn't have to be the one that uses it.  A common
convention in writing code with deferred locks is to acquire locks on behalf of a task being
launched.  Figure~\ref{fig:lockevents} illustrates this with a simple example.  The {\tt lock\_example}
function needs to launch two subtasks on two other processors, each with exclusive access to some resource protected by a
lock named {\tt mutex}.  For each subtask, it first requests the lock on the subtask's behalf.  Using the
event returned by the lock request as a precondition, it then spawns the subtask.  The runtime will guarantee
that the subtask is not started until the lock has been acquired.  The parent task also requests that the
runtime automatically release the lock after each subtask has completed by using the subtask's completion
event as a precondition for the unlock.  In the event dependency graph (Figure~\ref{fig:lockevents}),
solid arrows represent
the explicit events used in the code, while the dotted line represents the dynamic ordering that results
from the mutual exclusion guarantee of the lock.  (In this example, the second subtask is granted the lock first.)

@@

\subsection{Physical Regions}
\label{subsec:phyreg}
%In distributed machines with discrete memories, data movement often consists of more
%than a simple copy operation.  Many applications perform operations and transformations on
%their data in conjunction with data movement for higher performance.  One common example is 
%that applications accumulate reduction operations in a buffer and then, as part of a copy operation,
%the reduction operations are applied to a destination buffer on the receiving side of the copy.  
%To support these kinds of conjoined operations in an asynchronous environment, our interface 
%must be aware of the structure of data so it can perform accompanying operations with 
%data movement operations.  To describe the structure of data our interface uses {\em physical regions.}
Current APIs for moving data in machines with discrete memories only support untyped
copies of a buffer of data and force the programmer to work with pointers in a variety of address
spaces.  This restriction requires programmers to manage both the layout of
data and the marshaling of data for communication.  {\em Physical regions} are a
mechanism for decoupling the placement of data from its layout.  Physical regions allow portability
of ``pointers'' as data is moved around the memory hierarchy, and enable multiple data
layouts, such as specialized layouts for buffering individual reduction operations.

@@

%The data held inside a physical region is accessed by another kind of object called an
%{\em accessor} (not shown).  Accessors support read, write, and reduction operations
%to elements inside the physical region.  Accessors can be specialized for the case where
%all memory is known to be directly accessible to the executing processor, which allows 
%for direct references to elements.  In cases where not all memory is directly accessable, 
%accessors supply the level of indirection necessary to convert operations into remote 
%memory operations (RDMAs) if supported by the underlying hardware.  
%An example of this is described in more detail in Section~\ref{sec:impl}.

### 2012-11-14 doc/llr-pldi2013/micro.tex
Lots of small edits
last-minute edits, spell check

Our second micro-benchmark for events measures the maximum rate at which events can
be triggered by our implementation.  Instead of a single chain, many parallel chains
are created.  Additionally, each step of the chain consists of a parameterized number
of events, each one dependent on every event from the previous step of the chain.
This parameter is called the {\em fan-in/out factor}.  The events within each step of
a chain are distributed across the nodes.  Recall from Section~\ref{subsec:eventimpl}
that the aggregation of event subscriptions
will limit the number of event trigger active messages to one per node (per event trigger)
even when the fan-in/out factor exceeds the node count.

@@

Our lock micro-benchmark is designed to measure the rate at which deferred locks can be granted
in a distributed environment.
A parameterized number of locks are created per node and their handles are made available to
every node.  Each node then creates a parameterized number of {\em chains} of lock/unlock
request pairs, where each request pair is to a random lock, and is made dependent on the 
previous lock/unlock pair's completion by the event returned from the lock request.  
%The chains are made long enough that start-up
%and cleanup times are negligible.  
The total number of chains across all nodes gives the total
number of lock requests that can exist in the system at any given time.  All chains are
started at the same time (via a dependence on a single user event), and the total time to
process all the chains is divided into the total number of lock requests to yield an average
lock grant rate.

@@

Figure~\ref{fig:fixedlock} shows the lock grant rate for a variety of node counts and locks
per node.  The number of chains per node is varied so that the total number of chains in the system
is 1024 in all cases.  For the single-node cases, the insensitivity to the number of locks indicates
that the bottleneck is in the computational ability of the node to process the requests. 
For larger numbers of nodes, especially for the larger numbers of locks per node (where contention
for any given lock is low), the speedup with increasing node count suggests the limiting factor is
the rate at which lock-related active messages can be sent (nearly every request will require a
transfer of ownership).  Especially in the middle of the node count range, the performance actually
increases with
decreasing lock counts.  Although contention increases, the favoring of local lock requestors makes
the contention an advantage, reducing the number of lock-related active messages that must be sent
per lock grant.

@@

To more clearly demonstrate the benefit of lock unfairness, even at large node counts, we show an alternative
cut of the experiment space in Figure~\ref{fig:fixednode}.
In this case the node count is fixed at 8 and lock grant rates are shown for a variety of total lock counts
and number of chains per node.
With only 32 chains per node (a total of 256 chains)
there is little contention and the grant rate is high.  As the number of chains per node increases there is
more contention for locks and the rate drops as a result.  For smaller lock counts, further increases in
the number of chains result in improved rates.  Note that the point on each line where the increase occurs
is when the chains per node is greater than total locks, which corresponds to the point where the expected 
number of requests per lock 
per node becomes more than one.  As soon as there are multiple expected requestors for the same lock on
a node, the unfairness property of our lock implementation is able to reduce the number of lock ownership 
transfers, yielding better performance.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.33]{figs/reduce_simple.pdf}
\end{center}
\vspace{-2mm}
\caption{Reduction Benchmark.\label{fig:reducbench}}
\vspace{-4mm}
\end{figure}

>>>>

Our second micro-benchmark for events measures the maximum rate at which events can
be triggered by our implementation.  Instead of a single chain, many parallel chains
are created.  Additionally, each step of the chain consists of a parametrized number
of events, each one dependent on every event from the previous step of the chain.
This parameter is called the {\em fan-in/out factor}.  The events within each step of
a chain are distributed across the nodes.  Recall from Section~\ref{subsec:eventimpl}
that the aggregation of event subscriptions
will limit the number of event trigger active messages to one per node (per event trigger)
even when the fan-in/out factor exceeds the node count.

@@

Our lock micro-benchmark is designed to measure the rate at which deferred locks can be granted
in a distributed environment.
A parametrized number of locks are created per node and their handles are made available to
every node.  Each node then creates a parametrized number of {\em chains} of lock/unlock
request pairs, where each request pair is to a random lock, and is made dependent on the 
previous lock/unlock pair's completion by the event returned from the lock request.  
%The chains are made long enough that start-up
%and cleanup times are negligible.  
The total number of chains across all nodes gives the total
number of lock requests that can exist in the system at any given time.  All chains are
started at the same time (via a dependence on a single user event), and the total time to
process all the chains is divided into the total number of lock requests to yield an average
lock grant rate.

@@

Figure~\ref{fig:fixedlock} shows the lock grant rate for a variety of node counts and locks
per node.  The number of chains per node is varied so that the total number of chains in the system
is 1024 in all cases.  For the single-node cases, the insensitivity to the number of locks indicates
that the bottleneck is in the computational ability of the node to process the requests. 
For larger numbers of nodes, especially for the larger numbers of locks per node (where contention
for any given lock is low), the speedup with increasing node count suggests the limiting factor is
the rate at which lock-related active messages can be sent (nearly every request will require a
transfer of ownership).  Especially in the middle of the node count range, the performance actually
increases with
decreasing lock counts.  Although contention increases, the favoring of local lock requesters makes
the contention an advantage, reducing the number of lock-related active messages that must be sent
per lock grant.

@@

To more clearly demonstrate the benefit of lock unfairness, even at large node counts, we show an alternative
cut of the experiment space in Figure~\ref{fig:fixednode}.
In this case the node count is fixed at 8 and lock grant rates are shown for a variety of total lock counts
and number of chains per node.
With only 32 chains per node (a total of 256 chains)
there is little contention and the grant rate is high.  As the number of chains per node increases there is
more contention for locks and the rate drops as a result.  For smaller lock counts, further increases in
the number of chains result in improved performance.  Note that the point on each line where the increase occurs
is when the chains per node is greater than total locks, which corresponds to the point where the expected 
number of requests per lock 
per node becomes more than one.  As soon as there are multiple expected requesters for the same lock on
a node, the unfairness property of our lock implementation is able to reduce the number of lock ownership 
transfers, yielding better performance.

@@
